---
title: "Macrofaunal Total Biomass"
author: "Chih-Lin Wei"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###) 
    toc_float: true
    number_sections: true  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r}
library(bbbs)
library(ggplot2)
library(plyr)
library(doBy)
library(readxl)
library(knitr)
library(swtmap)
library(TWBathyMap)
library(maptools)
library(nlme)

large <- theme(legend.title = element_text(size=20),
        legend.text = element_text(size=20),
        axis.title = element_text(size=20),
        axis.text = element_text(size=20))

# dark theme for ggplot
dark <- theme(plot.background = element_rect(colour = 'NA', fill = 'gray10'),
        panel.background = element_rect(colour = 'NA', fill = 'transparent'),
        legend.background = element_rect(colour = 'NA', fill = 'transparent'),
        legend.title = element_text(colour = 'white', size=20),
        legend.text = element_text(colour = 'white', size=20),
        axis.title = element_text(colour = 'white', size=20),
        axis.text = element_text(colour = 'white', size=20),
        axis.ticks = element_line(colour = 'white'),
        panel.border = element_rect(fill = 'NA', colour = 'white'), 
        panel.grid.major = element_line(colour = 'gray30'),
        panel.grid.minor = element_line(colour = 'gray20'))

jet.colors <-colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan",
                     "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))
```

# Mean density 
## per cruise


```{r, fig.height=5, fig.width=7}
# Convert the numbers of polychaete specimen (with head) to abundance
pols <- subset(mac, Taxon=="Polychaeta" & (Condition=="FH" | Condition=="C" | Condition=="FHT"))
pola <- summaryBy(Size~Cruise+Station+Deployment+Tube+Taxon, data=pols, FUN=length, var.names="Abundance", keep.names = T)
pola <- cbind(pola[, 1:4], Section="0-15", pola[, 5:6], Comment="")

# Replace polycahete sorting abundance with correct abundance from size data
sort <- read_excel("../excel/GPSC_macro_sorting_2020.07.31.xlsx", sheet=2)
sort <- rbind(subset(sort, Taxon!="Polychaeta"), pola)
# Remove Cyclopoida and Calanoida
sort <- subset(sort, !(Taxon=="Cyclopoida"|Taxon=="Calanoida"))
# Total abundance by tube
abun <- summaryBy(Abundance~Cruise+Station+Deployment+Tube, data=as.data.frame(sort), FUN=sum, keep.names = TRUE)
depth <- read_excel("../excel/GPSC_macro_sorting_2020.07.31.xlsx", sheet=1)
abun <- cbind(depth[match(paste(abun$Cruise, abun$Station), paste(depth$Cruise, depth$Station)),], Abundance=abun$Abundance)


dat <- subset(abun, Cruise=="OR1_1096"|Cruise=="OR1_1102"|Cruise=="OR1_1114"|Cruise=="OR1_1126")
# Remove outlier samples with only 1 individuals
bad <- with(dat, Abundance <=1 | (Cruise=="OR1_1102" & Station == "GS4") )
dat <- subset(dat, !bad)

dat$Density <- dat$Abundance/dat$Area
dat.mean <- summaryBy(Density+Longitude+Latitude+Depth+Date~Habitat+Cruise+Station, data=dat, FUN=c(mean, sd, length))
dat.mean$Date.mean <- strftime(as.POSIXct(dat.mean$Date.mean, origin = "1970-01-01", tz = "Asia/Taipei"), format="%Y-%m-%d")
dat.mean$Density.se <- dat.mean$Density.sd/dat.mean$Density.length^0.5

out <- dat.mean[, c("Habitat", "Cruise", "Station", "Longitude.mean", "Latitude.mean", "Depth.mean", "Date.mean", "Density.length", "Density.mean", "Density.sd", "Density.se")]
names(out) <- c("Habitat", "Cruise", "Station", "Longitude", "Latitude", "Depth", "Date", "n", "Density", "sd", "se")

#out$Longitude <- round(out$Longitude, 4)
#out$Latitude <- round(out$Latitude, 4)
#out$Depth <- round(out$Depth, 0)
#out$Density <- round(out$Density, 0)
#out$sd <- round(out$sd, 1)
#out$se <- round(out$se, 1)
kable(out, align="c", caption="OR1_1102 GS4, OR1_1126 GC2, and OR1_1126 GC3 Tube 2 had very low abundance and only single replication. They are removed as outliers before plotting")
```

### Linear regression with depth

```{r}
splitBy(~Habitat, out) %>% lapply(FUN=function(x)lm(log10(Density)~Depth, data=x) %>% summary)

ggplot(data=out, 
       aes(x=Depth, y=log10(Density), 
           ymin=log10(Density-se), ymax=log10(Density+se),
           colour=Habitat, fill=Habitat, linetype=Habitat,
           shape=Habitat))+
  geom_point(size=5, position = "dodge")+ 
  stat_smooth(data=subset(out, Habitat=="Canyon"), method="lm", formula=y~x, alpha=0.2)+
  geom_errorbar(linetype=1, position = "dodge")+
  scale_colour_manual(values=c("black","gray30"))+
  scale_fill_manual(values=c("black","gray30"))+
  scale_linetype_manual(values=c(1,2))+
  scale_shape_manual(values=c(19,1))+
  labs(x="Depth (m)", y=expression(Log[10]~density~(ind~m^-2)))+
  #scale_y_log10()+
  theme_bw() %+replace% large #%+replace% dark
```

Significant relationship between macrofauna density and water depth in the canyon.

### Generalized Least Squares (GLS) modeling

Macrofauna density of each sample were averaged by station (i.e., average of three cores) before further statistical tests. We used Generalized Least Squares (GLS) modeling to examine the effects of habitat (canyon vs. slope), depth, and sampling time on the mean macrofauna density.

#### Main test

```{r}
f <- gls(log10(Density) ~ Habitat+Depth+Cruise+Habitat:Depth+Habitat:Cruise+Depth:Cruise, data=out, method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
```

Significant habitat and depth effects, but there are also significant interactions between habitat and depth, as well as between habitat and cruise.

```{r, fig.width=12, fig.height=12}
# Function to plot dianotics plot
dianostic_plot <- 
  function(f, y){
    # standardized residuals versus fitted values
    a1 <- plot(f, resid(., type = "p") ~ fitted(.) | Habitat, abline = 0)
    a2 <- plot(f, resid(., type = "p") ~ fitted(.) | Cruise, abline = 0)
    a3 <- plot(f, resid(., type = "p") ~ fitted(.), abline = 0)
    # box-plots of residuals
    b1<-plot(f, Habitat ~ resid(.))
    b2 <- plot(f, Cruise ~ resid(.))
    # observed versus fitted values
    c1<-plot(f, paste(paste(y, "fitted(.)", sep="~"), "Habitat", sep="|") %>% formula, abline = c(0,1))
    c2<-plot(f, paste(paste(y, "fitted(.)", sep="~"), "Cruise", sep="|") %>% formula, abline = c(0,1))
    c3<-plot(f, paste(y, "fitted(.)", sep="~") %>% formula, abline = c(0,1))
    # QQ plot
    d1<-qqnorm(f, ~ resid(., type = "p") | Habitat, abline = c(0,1))
    d2<-qqnorm(f, ~ resid(., type = "p") | Cruise, abline = c(0,1))
    d3<-qqnorm(f, ~ resid(., type = "p"), abline = c(0,1))
    
    print(a1, split=c(1,1,3,4), more=TRUE)
    print(a2, split=c(2,1,3,4), more=TRUE)
    print(a3, split=c(3,1,3,4), more=TRUE)
    print(b1, split=c(1,2,3,4), more=TRUE)
    print(b2, split=c(2,2,3,4), more=TRUE)
    #
    print(c1, split=c(1,3,3,4), more=TRUE)
    print(c2, split=c(2,3,3,4), more=TRUE)
    print(c3, split=c(3,3,3,4), more=TRUE)
    print(d1, split=c(1,4,3,4), more=TRUE)
    print(d2, split=c(2,4,3,4), more=TRUE)
    print(d3, split=c(3,4,3,4))
    }

dianostic_plot(f, y = "log10(Density)")
```

#### Pairwise tests

```{r}
# In canyon
f <- gls(log10(Density) ~ Depth+Cruise+Depth:Cruise, data=subset(out, Habitat == "Canyon"), method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
# On slope
f <- gls(log10(Density) ~ Depth+Cruise+Depth:Cruise, data=subset(out, Habitat == "Slope"), method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
# OR1 1096
f <- gls(log10(Density) ~ Habitat+Depth+Habitat:Depth, data=subset(out, Cruise=="OR1_1096"), method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
# OR1 1102
f <- gls(log10(Density) ~ Habitat+Depth+Habitat:Depth, data=subset(out, Cruise=="OR1_1102"), method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
# OR1 1114
f <- gls(log10(Density) ~ Habitat+Depth+Habitat:Depth, data=subset(out, Cruise=="OR1_1114"), method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
# OR1 1126
f <- gls(log10(Density) ~ Habitat+Depth+Habitat:Depth, data=subset(out, Cruise=="OR1_1126"), method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
```

Significant depth effect in the canyon; significant habitat effects during OR1 1102, 1114, and 1126.

## Per station

```{r, fig.height=5, fig.width=7}
dat.mean <- summaryBy(Density+Longitude+Latitude+Depth~Habitat+Station, data=out, FUN=c(mean, sd, length))
dat.mean$Density.se <- dat.mean$Density.sd/dat.mean$Density.length^0.5

out <- dat.mean[,c("Habitat", "Station", "Longitude.mean", "Latitude.mean", "Depth.mean", "Density.length", "Density.mean", "Density.sd", "Density.se")]
names(out) <- c("Habitat", "Station", "Longitude", "Latitude", "Depth","n", "Density", "sd", "se")

out$Longitude <- round(out$Longitude, 4)
out$Latitude <- round(out$Latitude, 4)
out$Depth <- round(out$Depth, 0)
out$Density <- round(out$Density, 0)
out$sd <- round(out$sd, 1)
out$sd <- round(out$se, 1)
kable(out, align="c")

ggplot(data=out, 
       aes(x=Depth, y=log10(Density), 
           ymin=log10(Density-se), ymax=log10(Density+se),
           colour=Habitat, fill=Habitat, linetype=Habitat, shape=Habitat))+
  geom_point(size=5, position = "dodge")+ 
  stat_smooth(method="lm", formula=y~x, alpha=0.2)+
  geom_errorbar(linetype=1, position = "dodge")+
  scale_colour_manual(values=c("black","gray30"))+
  scale_fill_manual(values=c("black","gray30"))+
  scale_linetype_manual(values=c(1,2))+
  scale_shape_manual(values=c(19,1))+
  labs(x="Depth (m)", y=expression(Log[10]~density~(ind~m^-2)))+
  theme_bw() %+replace% large #%+replace% dark
```

# Mean biomass 
## Per cruise

```{r, fig.height=5, fig.width=7}
dat <- mac

dat <- subset(dat, Cruise=="OR1_1096"|Cruise=="OR1_1102"|Cruise=="OR1_1114"|Cruise=="OR1_1126")

dat.sum <- summaryBy(Wt+Area+Longitude+Latitude+Depth~Habitat+Cruise+Station+Deployment+Tube, data=dat, FUN=c(sum, mean))
dat.sum$Biomass <- dat.sum$Wt.sum/dat.sum$Area.mean

dat.mean <- summaryBy(Biomass+Longitude.mean+Latitude.mean+Depth.mean~Habitat+Cruise+Station, data=dat.sum, FUN=c(mean, sd, length))
dat.mean$Biomass.se <- dat.mean$Biomass.sd/dat.mean$Biomass.length^0.5

out <- dat.mean[, c("Habitat", "Cruise", "Station", "Longitude.mean.mean", "Latitude.mean.mean", "Depth.mean.mean", "Biomass.length", "Biomass.mean", "Biomass.sd", "Biomass.se")]
names(out) <- c("Habitat", "Cruise", "Station", "Longitude", "Latitude", "Depth", "n", "Biomass", "sd", "se")

#out$Longitude <- round(out$Longitude, 4)
#out$Latitude <- round(out$Latitude, 4)
#out$Depth <- round(out$Depth, 0)
#out$Biomass <- round(out$Biomass, 0)
#out$sd <- round(out$sd, 1)
#out$se <- round(out$se, 1)
kable(out, align="c")
```

### Linear regression with depth

```{r}
splitBy(~Habitat, out) %>% lapply(FUN=function(x)lm(log10(Biomass)~Depth, data=x) %>% summary)

ggplot(data=out, 
       aes(x=Depth, y=log10(Biomass), 
           ymin=log10(Biomass-se), ymax=log10(Biomass+se),
           shape=Habitat, colour=Habitat, fill=Habitat, linetype=Habitat))+
  geom_point(size=5, position = "dodge")+
  stat_smooth(data=subset(out, Habitat=="Canyon"), method="lm", formula=y~x, alpha=0.2)+
  geom_errorbar(linetype=1, position = "dodge")+
  scale_colour_manual(values=c("black","gray30"))+
  scale_fill_manual(values=c("black","gray"))+
  scale_linetype_manual(values=c(1,2))+
  scale_shape_manual(values=c(19,1))+
  labs(x="Depth (m)", y=expression(Log[10]~biomass~(mg~m^-2)))+
  #scale_y_log10()+
  theme_bw() %+replace% large #%+replace% dark
```

Marginal relationship between macrofauna biomass and water depth in the canyon.

### Generalized Least Squares (GLS) modeling

Macrofauna biomass of each sample were averaged by station (i.e., average of three cores) before further statistical tests. We used Generalized Least Squares (GLS) modeling to examine the effects of habitat (canyon vs. slope), depth, and sampling time on the mean macrofauna Biomass.

#### Main test

```{r}
f <- gls(log10(Biomass) ~ Habitat+Depth+Cruise+Habitat:Depth+Habitat:Cruise+Depth:Cruise, data=out, method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
```

Significant habitat effect, but there are also significant interactions between habitat and depth.

```{r, fig.width=12, fig.height=12}
# Dianotics plot
dianostic_plot(f, y = "log10(Biomass)")
```

#### Pairwise tests

```{r}
# In canyon
f <- gls(log10(Biomass) ~ Depth+Cruise+Depth:Cruise, data=subset(out, Habitat == "Canyon"), method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
# On slope
f <- gls(log10(Biomass) ~ Depth+Cruise+Depth:Cruise, data=subset(out, Habitat == "Slope"), method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
```

Significant depth effect on slope

## Per station

```{r, fig.height=5, fig.width=7}
dat.mean <- summaryBy(Biomass+Longitude+Latitude+Depth~Habitat+Station, data=out, FUN=c(mean, sd, length))
dat.mean$Biomass.se <- dat.mean$Biomass.sd/dat.mean$Biomass.length^0.5
out <- dat.mean[, c("Habitat", "Station", "Longitude.mean", "Latitude.mean", "Depth.mean", "Biomass.length", "Biomass.mean", "Biomass.sd", "Biomass.se")]
names(out) <- c("Habitat", "Station", "Longitude", "Latitude", "Depth", "n", "Biomass", "sd", "se")

out$Longitude <- round(out$Longitude, 4)
out$Latitude <- round(out$Latitude, 4)
out$Depth <- round(out$Depth, 0)
out$Biomass <- round(out$Biomass, 0)
out$sd <- round(out$sd, 1)
out$se <- round(out$se, 1)
kable(out, align="c")

ggplot(data=out, 
       aes(x=Depth, y=log10(Biomass), 
           ymin=log10(Biomass-se), ymax=log10(Biomass+se),
           shape=Habitat, colour=Habitat, fill=Habitat, linetype=Habitat))+
  geom_point(size=5, position = "dodge")+
  stat_smooth(method="lm", formula=y~x, alpha=0.2)+
  scale_colour_manual(values=c("black","gray30"))+
  scale_fill_manual(values=c("black","gray30"))+
  scale_shape_manual(values=c(19,1))+
  geom_errorbar(linetype=1, position = "dodge")+
  scale_linetype_manual(values=c(1,2))+
  labs(x="Depth (m)", y=expression(Log[10]~biomass~(mg~m^-2)))+
  theme_bw() %+replace% large #%+replace% dark
```

# Sampling map

```{r, fig.height=8, fig.width=8}
tbspplot <- 
  function(x, ...){
  slope <- terrain(x, opt="slope")
  aspect <- terrain(x, opt="aspect")
  hill <- hillShade(slope, aspect, 80, 270)
  p2 <- spplot(hill, col.regions=grey(0:100/100), cut=100, colorkey=FALSE)
  pal <- tb.colors(x)
  p1 <- spplot(x, col.regions=alpha(pal$col, 0.6), at=pal$breaks,
               colorkey=list(space="top", labels=list(cex=1)),
               scales=list(draw = TRUE, cex=1)
               ,...)
p1+as.layer(p2, under=T)
}


loc <- out
coordinates(loc) <- c("Longitude", "Latitude")
r <- crop(swtmap::bathy, extent(120.15, 120.45, 22.1, 22.5))
#r <- projectRaster(r, res=0.001, crs="+proj=longlat")
isobath<- rasterToContour(r, levels=c(seq(-1800, -200, by=100), 0))

extra <- function(...){
         panel.levelplot.raster(...)
         sp.lines(isobath, col="gray30")
         sp.points(loc, cex=2, pch=19, col=rep(c("#e41a1c", "#4daf4a"), each=4))
         sp.pointLabel(loc, labels=loc$Station, cex=1.5, allowSmallOverlap=FALSE)
         SpatialPolygonsRescale(layout.scale.bar(), offset=c(120.28, 22.46), scale = 10/(111.321*cos(22.46*pi/180)), fill = c("white","black"), col = "black")
         sp.text(loc = c(120.28, 22.475), "0")
         sp.text(loc = c(120.38, 22.475), "10 km")
}
tbspplot(r, panel=extra)
```

```{r, fig.height=8, fig.width=8}
e <- extent(119.1, 122.2, 21.6, 25.5)
r <- crop(TWBathyMap::bathy, e)
#r <- projectRaster(r, res=0.001, crs="+proj=longlat")
lon <- c(120.15, 120.15, 120.45, 120.45, 120.15)
lat <- c(22.1, 22.5, 22.5, 22.1, 22.1)
b <- SpatialPolygons(list(Polygons(list(Polygon(cbind(lon, lat))), 1)))
extra <- function(...){
         panel.levelplot.raster(...)
         sp.polygons(b, col="#e41a1c", lty=2)
         sp.points(loc, cex=0.5, pch=19, col=rep(c("#e41a1c", "#4daf4a"), each=4))
}
tbspplot(r, panel=extra)
```

