---
title: "Macrofauna Respiration"
author: "Chih-Lin Wei"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###) 
    toc_float: true
    number_sections: true  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(bbbs)
library(BenthicPro)
library(ggplot2)
library(plyr)
library(doBy)
library(readxl)
library(knitr)
library(nlme)

large <- theme(legend.title = element_text(size=20),
        legend.text = element_text(size=20),
        axis.title = element_text(size=20),
        axis.text = element_text(size=20))
rotate <- theme(axis.text.x = element_text(size=20, angle=60, hjust=0.5))

# dark theme for ggplot
dark <- theme(plot.background = element_rect(colour = 'NA', fill = 'gray10'),
        panel.background = element_rect(colour = 'NA', fill = 'transparent'),
        legend.background = element_rect(colour = 'NA', fill = 'transparent'),
        legend.title = element_text(colour = 'white', size=20),
        legend.text = element_text(colour = 'white', size=20),
        axis.title = element_text(colour = 'white', size=20),
        axis.text = element_text(colour = 'white', size=20),
        axis.ticks = element_line(colour = 'white'),
        panel.border = element_rect(fill = 'NA', colour = 'white'), 
        panel.grid.major = element_line(colour = 'gray30'),
        panel.grid.minor = element_line(colour = 'gray20'))
```

# Data preparation

Mass-specific macrofauna respiration rate was estimated from body mass and temperature using an empirical model, $log(R/M) = 8.3732 - 0.2073 * log(M) - 2766.0235 / T$, based on  a large database compilated for aquatic invertebrate respriation (<22000 measurements, >900 species, > 440 references) by [Thomas Brey](http://www.thomas-brey.de/science/virtualhandbook/respir/rempirics1.html). In the equation, R/M is the mass-specific respiration (J/J/day), M is individual body mass (J), and T is water temperature (K). 

```{r}
dat <- subset(mac, Cruise=="OR1_1096"|Cruise=="OR1_1102"|Cruise=="OR1_1114"|Cruise=="OR1_1126"|Cruise=="OR1_1128"|Cruise=="OR1_1132")
dat.sum <- summaryBy(Wt~Habitat+Cruise+Station+Deployment+Tube+Taxon, data=dat, FUN=c(sum, length))
names(dat.sum)[-1:-6] <- c("Wt", "Count")
dat.sum$Sample.area <- mac$Area[match(with(dat.sum, paste(Cruise, Station)), with(mac, paste(Cruise, Station)))]

# Convert the numbers of polychaete specimen (with head) to abundance
pols <- subset(mac, Taxon=="Polychaeta" & (Condition=="FH" | Condition=="C" | Condition=="FHT"))
pola <- summaryBy(Wt~Habitat+Cruise+Station+Deployment+Tube+Taxon, data=pols, FUN=length, var.names="Count", keep.names = T)
keep <- match(with(pola, paste(Cruise, Station, Deployment, Tube, Taxon)), with(dat.sum, paste(Cruise, Station, Deployment, Tube, Taxon)))
dat.sum$Count[keep] <- pola$Count

# Convert the numbers of complete ophiroid specimens to abundance
ophs <- subset(mac, Taxon=="Ophiuroidea" & Condition=="C")
opha <- summaryBy(Wt~Habitat+Cruise+Station+Deployment+Tube+Taxon, data=ophs, FUN=length, var.names="Count", keep.names = T)
keep <- match(with(opha, paste(Cruise, Station, Deployment, Tube, Taxon)), with(dat.sum, paste(Cruise, Station, Deployment, Tube, Taxon)))
dat.sum$Count[keep] <- opha$Count

# Remove Cyclopoida and Calanoida
dat.sum <- subset(dat.sum, !(Taxon=="Cyclopoida"|Taxon=="Calanoida"))

# Match to marofauna respiration parameters
dat.sum <- cbind(dat.sum, mpr[match(dat.sum$Taxon, mpr$Taxon), c("ConFac_j2mgwm", "ConFac_j2mgc")])
dat.sum$Bodymass <- with(dat.sum, Wt*ConFac_j2mgwm/Count)
dat.sum <- cbind(dat.sum, env[match(with(dat.sum, paste(Cruise, Station)), with(env, paste(Cruise, Station))), c("Temperature", "Depth")])
dat.sum <- cbind(dat.sum, mpr[match(dat.sum$Taxon, mpr$Taxon), c(-1:-3, -21)])

# Calculate R/M ratio
# Convert bodymass in J to kcal
individualRM <- 8.3732-0.2073*log10(dat.sum$Bodymass)-2766.0235 /(dat.sum$Temperature+273.15)

RMresults<-cbind(dat.sum, daily.RtoM=10^individualRM)

# Daily respiration in mg WM per species and sample per square meter
RMresults$respirationW <- with(RMresults, daily.RtoM*Wt/Sample.area)

# Daily respiration in J per species and sample per square meter
RMresults$respirationJ <- with(RMresults, respirationW*ConFac_j2mgwm)

# Daily respiration in milligram carbon per species and sample per square meter
RMresults$respirationC <- with(RMresults, respirationJ/ConFac_j2mgc)


# Average P/B ratio by tube
p2b <- summaryBy(Depth+daily.RtoM~Habitat+Cruise+Station+Deployment+Tube, data = RMresults, fun = mean, keep.names = TRUE)

# Summation production by tube
pd <- summaryBy(respirationW+respirationJ+respirationC+Count+Wt~Habitat+Cruise+Station+Deployment+Tube, data = RMresults, fun = sum, keep.names = TRUE)
pd <- cbind(p2b, pd[, -1:-5])

se <- function(x) sd(x)/length(x)^0.5
out <- summaryBy(Depth+daily.RtoM+respirationW+respirationJ+respirationC~Habitat+Cruise+Station, data = pd, FUN = c(mean, sd, length, se))

out <- out[, c("Habitat", "Cruise", "Station", "Depth.mean", "daily.RtoM.mean", "daily.RtoM.se", "respirationW.mean", "respirationW.se", "respirationJ.mean", "respirationJ.se", "respirationC.mean", "respirationC.se", "Depth.length")]

names(out) <- c("Habitat", "Cruise", "Station", "Depth", "daily.RtoM", "daily.RtoM.se", "respirationW", "respirationW.se", "respirationJ", "respirationJ.se", "respirationC", "respirationC.se", "n")

kable(out)
```
# Mass-specific respiration rate
## Linear regression with depth

```{r, fig.height=5, fig.width=7}
splitBy(~Habitat, out) %>% lapply(FUN=function(x)lm(daily.RtoM~Depth, data=x) %>% summary)

(f5b <- ggplot(data=out, 
       aes(x=Depth, y=daily.RtoM, 
           ymin=daily.RtoM-daily.RtoM.se, ymax=daily.RtoM+daily.RtoM.se,
           shape=Habitat, colour=Habitat, fill=Habitat, linetype=Habitat))+
  geom_point(size=5, position="jitter")+
  stat_smooth(method="lm", formula=y~x, alpha=0.2, size=1.5)+
  geom_errorbar(linetype=1, position="jitter")+
  scale_colour_manual(values=c("#e41a1c", "#4daf4a"))+
  scale_fill_manual(values=c("#e41a1c", "#4daf4a"))+
  scale_linetype_manual(values=c(1,2))+
  scale_shape_manual(values=c(19,1))+
  labs(x="Depth (m)", y=expression("Mass-specific"~respiration~(day^-1)))+
  theme_bw() %+replace% large #%+replace% dark
)

save(f5b, file="../rds/f5b.rds")
```

Mass-specific respiration rate declined significantly with depth in the canyon and on the slope.

## Generalized Least Squares (GLS) modeling

Mass-specific respiration of each sample were averaged by core (i.e., average of three subcores) before further statistical tests. We used Generalized Least Squares (GLS) modeling to examine the effects of habitat (canyon vs. slope), depth, and sampling time on the mean Mass-specific respiration rate.

#### Main test

```{r}
f <- gls(daily.RtoM ~ Habitat+Depth+Cruise+Habitat:Depth+Habitat:Cruise+Depth:Cruise, data=out, method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
```

Significant depth effects.

```{r, fig.width=12, fig.height=12}
# Function to plot dianotics plot
dianostic_plot <- 
  function(f, y){
    # standardized residuals versus fitted values
    a1 <- plot(f, resid(., type = "p") ~ fitted(.) | Habitat, abline = 0)
    a2 <- plot(f, resid(., type = "p") ~ fitted(.) | Cruise, abline = 0)
    a3 <- plot(f, resid(., type = "p") ~ fitted(.), abline = 0)
    # box-plots of residuals
    b1<-plot(f, Habitat ~ resid(.))
    b2 <- plot(f, Cruise ~ resid(.))
    # observed versus fitted values
    c1<-plot(f, paste(paste(y, "fitted(.)", sep="~"), "Habitat", sep="|") %>% formula, abline = c(0,1))
    c2<-plot(f, paste(paste(y, "fitted(.)", sep="~"), "Cruise", sep="|") %>% formula, abline = c(0,1))
    c3<-plot(f, paste(y, "fitted(.)", sep="~") %>% formula, abline = c(0,1))
    # QQ plot
    d1<-qqnorm(f, ~ resid(., type = "p") | Habitat, abline = c(0,1))
    d2<-qqnorm(f, ~ resid(., type = "p") | Cruise, abline = c(0,1))
    d3<-qqnorm(f, ~ resid(., type = "p"), abline = c(0,1))
    
    print(a1, split=c(1,1,3,4), more=TRUE)
    print(a2, split=c(2,1,3,4), more=TRUE)
    print(a3, split=c(3,1,3,4), more=TRUE)
    print(b1, split=c(1,2,3,4), more=TRUE)
    print(b2, split=c(2,2,3,4), more=TRUE)
    #
    print(c1, split=c(1,3,3,4), more=TRUE)
    print(c2, split=c(2,3,3,4), more=TRUE)
    print(c3, split=c(3,3,3,4), more=TRUE)
    print(d1, split=c(1,4,3,4), more=TRUE)
    print(d2, split=c(2,4,3,4), more=TRUE)
    print(d3, split=c(3,4,3,4))
    }

dianostic_plot(f, y = "daily.RtoM")
```

# Daily respiration in wet weight per square meter
## Linear regression with depth

```{r, fig.height=5, fig.width=7}
splitBy(~Habitat, out) %>% lapply(FUN=function(x)lm(log10(respirationW)~Depth, data=x) %>% summary)

(f5d <- ggplot(data=out, 
       aes(x=Depth, y=log10(respirationW), 
           ymin=log10(respirationW-respirationW.se), ymax=log10(respirationW+respirationW.se),
           shape=Habitat, colour=Habitat, fill=Habitat, linetype=Habitat))+
  geom_point(size=5, position="jitter")+
  stat_smooth(data=subset(out, Habitat=="Slope"), method="lm", formula=y~x, alpha=0.2, size=1.5)+
  geom_errorbar(linetype=1, position="jitter")+
  scale_colour_manual(values=c("#e41a1c", "#4daf4a"))+
  scale_fill_manual(values=c("#e41a1c", "#4daf4a"))+
  scale_linetype_manual(values=c(1,2))+
  scale_shape_manual(values=c(19,1))+
  labs(x="Depth (m)", y=expression(Log[10]~respiration~(mg~m^-2~day^-1)))+
  theme_bw() %+replace% large #%+replace% dark
)

save(f5d, file="../rds/f5d.rds")
```

Respiration declined with depth on the slope. 

## Generalized Least Squares (GLS) modeling

Macrofauna respiration of each sample were averaged by station (i.e., average of three cores) before further statistical tests. We used Generalized Least Squares (GLS) modeling to examine the effects of habitat (canyon vs. slope), depth, and sampling time on the mean macrofauna respiration.

## Main test

```{r}
f <- gls(log10(respirationW) ~ Habitat+Depth+Cruise+Habitat:Depth+Habitat:Cruise+Depth:Cruise, data=out, method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
```

Significant habitat and depth effect, but there are also a significant interaction between habitat and depth.

```{r, fig.width=12, fig.height=12}
# Dianotics plot
dianostic_plot(f, y = "log10(respirationW)")
```

## Pairwise tests

```{r}
# In canyon
f <- gls(log10(respirationW) ~ Depth+Cruise+Depth:Cruise, data=subset(out, Habitat == "Canyon"), method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
# On slope
f <- gls(log10(respirationW) ~ Depth+Cruise+Depth:Cruise, data=subset(out, Habitat == "Slope"), method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
# GC1 and GS1
f <- gls(log10(respirationW) ~ Habitat, data=subset(out, Station=="GC1"|Station=="GS1"), method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
# GC2 and GS2
f <- gls(log10(respirationW) ~ Habitat, data=subset(out, Station=="GC2"|Station=="GS2"), method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
# GC3 and GS3
f <- gls(log10(respirationW) ~ Habitat, data=subset(out, Station=="GC3"|Station=="GS3"), method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
# GC4 and GS4
f <- gls(log10(respirationW) ~ Habitat, data=subset(out, Station=="GC4"|Station=="GS4"), method = "REML")
kable(summary(f)$tTable)
kable(anova(f))
```

Only significant depth effect on slope

# Daily respiration in carbon per square meter
## Linear regression with depth

```{r, fig.height=5, fig.width=7}
splitBy(~Habitat, out) %>% lapply(FUN=function(x)lm(log10(respirationC)~Depth, data=x) %>% summary)

ggplot(data=out, 
       aes(x=Depth, y=log10(respirationC), 
           ymin=log10(respirationC-respirationC.se), ymax=log10(respirationC+respirationC.se),
           shape=Habitat, colour=Habitat, fill=Habitat, linetype=Habitat))+
  geom_point(size=5, position="jitter")+
  stat_smooth(data=subset(out, Habitat=="Slope"), method="lm", formula=y~x, alpha=0.2, size=1.5)+
  geom_errorbar(linetype=1, position="jitter")+
  scale_colour_manual(values=c("#e41a1c", "#4daf4a"))+
  scale_fill_manual(values=c("#e41a1c", "#4daf4a"))+
  scale_linetype_manual(values=c(1,2))+
  scale_shape_manual(values=c(19,1))+
  labs(x="Depth (m)", y=expression(Log[10]~respiration~(mgC~m^-2~day^-1)))+
  theme_bw() %+replace% large #%+replace% dark
```

Respiration declined with depth on the slope. 

# Daily respiration in J per square meter
## Linear regression with depth

```{r, fig.height=5, fig.width=7}
splitBy(~Habitat, out) %>% lapply(FUN=function(x)lm(log10(respirationJ)~Depth, data=x) %>% summary)

ggplot(data=out, 
       aes(x=Depth, y=log10(respirationJ), 
           ymin=log10(respirationJ-respirationJ.se), ymax=log10(respirationJ+respirationJ.se),
           shape=Habitat, colour=Habitat, fill=Habitat, linetype=Habitat))+
  geom_point(size=5, position="jitter")+
  stat_smooth(data=subset(out, Habitat=="Slope"), method="lm", formula=y~x, alpha=0.2, size=1.5)+
  geom_errorbar(linetype=1, position="jitter")+
  scale_colour_manual(values=c("#e41a1c", "#4daf4a"))+
  scale_fill_manual(values=c("#e41a1c", "#4daf4a"))+
  scale_linetype_manual(values=c(1,2))+
  scale_shape_manual(values=c(19,1))+
  labs(x="Depth (m)", y=expression(Log[10]~respiration~(J~m^-2~yr^-1)))+
  theme_bw() %+replace% large #%+replace% dark
```

Respiration declined with depth on the slope. 
