---
title: "Meiofauna + Macrofauna Size Spectrum"
author: "Chih-Lin Wei"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(bbbs)
library(ggplot2)
library(plyr)
library(doBy)

large <- theme(legend.title = element_text(size=20),
        legend.text = element_text(size=20),
        strip.text=element_text(size=20),
        axis.title = element_text(size=20),
        axis.text.y = element_text(size=20),
        axis.text.x = element_text(size=20, angle=60, hjust=0.5))

rotate <- theme(axis.text.x = element_text(size=20, angle=60, hjust=0.5))

# dark theme for ggplot
dark <- theme(plot.background = element_rect(colour = 'NA', fill = 'gray10'),
        panel.background = element_rect(colour = 'NA', fill = 'transparent'),
        legend.background = element_rect(colour = 'NA', fill = 'transparent'),
        legend.title = element_text(colour = 'white', size=20),
        legend.text = element_text(colour = 'white', size=20),
        axis.title = element_text(colour = 'white', size=20),
        axis.text = element_text(colour = 'white', size=20),
        axis.ticks = element_line(colour = 'white'),
        panel.border = element_rect(fill = 'NA', colour = 'white'), 
        panel.grid.major = element_line(colour = 'gray30'),
        panel.grid.minor = element_line(colour = 'gray20'))
```


# Construction of meiofauna + macrofauna size specturm

Normalized biomass size spectrum (NBSS) and individual size distribution (ISD) were computed using custom R functions from Edwards et al. (2017, Methods in Ecology and Evolution 1:57-67). A probability density function based on a bounded power law (PLB) distribution was fitted to the ISD of meiofauna+macrofauna using maximum likelihood estimation (MLE) to estimate the exponent b. The exponent b (determined by MLE) is equivalent to the regression slope of the NBSS; however, comparison studies by Edwards et al. (2017) concluded that MLE method is the most accurate way to estimate the slope.
For normalized biomass size spectrum (NBSS), individual size of meiofauna+macrofauna were binned by log10 transformation of their biomass. The total biomass of each bin was normalized by the width of the bin. The log10 normalized biomass was plotted against the log10 midpoint of the size bin; however, the regression slope (and line) was not estimated from the NBSS. Instead, the slope (and line) was estimated by a customized probability density function using maximum likelihood.
The same fitted line (with 95% confidence interval) was plotted on the individual size distribution, which ranks the body size in decreasing order to visualize the fit between the model and size data.

# NBSS and ISD by habitat

```{r, fig.width=7, fig.height=5}
# Use functions from Edwards etal. (2017) Methods in Ecology and Evolution 1:57-67
# https://github.com/andrew-edwards/fitting-size-spectra
source("PLBfunctions.r")

hab <- splitBy(~Habitat, data=si)
wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=LBNbiom.method)
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- "Habitat"
nbss <- cbind(info, nbss[, -1])

#wt <- wt[unlist(lapply(wt, length))>9]
out <- lapply(wt, FUN=MLE.method)
# size specutrum slope estimate by MLE
#lapply(out, FUN=function(x)x[1:2])
plb <- ldply(lapply(out, FUN=function(x)x$PLB))
info <- ldply(strsplit(plb$.id, split="[|]"))
names(info) <- "Habitat"
plb <- cbind(info, plb[, -1])

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, shape=Habitat, colour=Habitat))+
  geom_point(size=3)+
  geom_path()+
  scale_shape_manual(values=c(19,1))+
  #geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB)))+
  #geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.LC)), linetype=2)+
  #geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.UC)), linetype=2)+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large %+replace% dark

<<<<<<< HEAD
# Remove sisze < 5e-4
=======
# Only use size > 1e-3
>>>>>>> 583f1dc7f3e655c5b4c6e206fc1beb89c8111286

hab <- splitBy(~Habitat, data=subset(si, Size>1e-3))
wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=LBNbiom.method)
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- "Habitat"
nbss <- cbind(info, nbss[, -1])

#wt <- wt[unlist(lapply(wt, length))>9]
out <- lapply(wt, FUN=MLE.method)
# size specutrum slope estimate by MLE
lapply(out, FUN=function(x)x[1:2])
plb <- ldply(lapply(out, FUN=function(x)x$PLB))
info <- ldply(strsplit(plb$.id, split="[|]"))
names(info) <- "Habitat"
plb <- cbind(info, plb[, -1])

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, shape=Habitat, colour=Habitat))+
  geom_point(size=3)+
  scale_shape_manual(values=c(19,1))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB)))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.LC)), linetype=2)+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.UC)), linetype=2)+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large %+replace% dark
```


```{r, fig.width=7, fig.height=5}
# Rank individual weight
SortWt <- lapply(hab, FUN=function(x)cbind(rank=1:length(x$Wt), weight=sort(x$Wt, decreasing=TRUE)))
SortWt <- ldply(SortWt)
info <- ldply(strsplit(SortWt$.id, split="[|]"))
names(info) <- "Habitat"
SortWt <- cbind(info, SortWt[, -1])

ggplot(data=SortWt, aes(x=weight, y=rank, shape=Habitat, colour=Habitat))+
  geom_point(size=2)+
  scale_shape_manual(values=c(19,1))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.LC), linetype=2)+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.UC), linetype=2)+
  #facet_wrap(~Cruise, scales="free")+
  scale_x_log10(expression(Individual~weight~(mg)), labels = function(x) format(x, scientific = T))+
  scale_y_log10("Individual rank", limits=c(1, NA), labels = function(x) format(x, scientific = T))+
  theme_bw() %+replace% large %+replace% dark
```

# NBSS and ISD by cruise and habitat

```{r, fig.width=7, fig.height=5}
hab <- splitBy(~Cruise+Habitat, data=si)
wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=LBNbiom.method)
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat")
nbss <- cbind(info, nbss[, -1])

#wt <- wt[unlist(lapply(wt, length))>9]
out <- lapply(wt, FUN=MLE.method)
# size specutrum slope estimate by MLE
#lapply(out, FUN=function(x)x[1:2])
plb <- ldply(lapply(out, FUN=function(x)x$PLB))
info <- ldply(strsplit(plb$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat")
plb <- cbind(info, plb[, -1])

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, shape=Habitat, colour=Habitat))+
  geom_point(size=3)+
  geom_path()+
  scale_shape_manual(values=c(19,1))+
  #geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB)))+
  #geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.LC)), linetype=2)+
  #geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.UC)), linetype=2)+
  facet_wrap(~Cruise)+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large %+replace% dark

# Only use size > 1e-3

hab <- splitBy(~Cruise+Habitat, data=subset(si, Size>1e-3))
wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=LBNbiom.method)
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat")
nbss <- cbind(info, nbss[, -1])

#wt <- wt[unlist(lapply(wt, length))>9]
out <- lapply(wt, FUN=MLE.method)
# size specutrum slope estimate by MLE
lapply(out, FUN=function(x)x[1:2])
plb <- ldply(lapply(out, FUN=function(x)x$PLB))
info <- ldply(strsplit(plb$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat")
plb <- cbind(info, plb[, -1])

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, shape=Habitat, colour=Habitat))+
  geom_point(size=3)+
  scale_shape_manual(values=c(19,1))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB)))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.LC)), linetype=2)+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.UC)), linetype=2)+
  facet_wrap(~Cruise)+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large %+replace% dark
```

```{r, fig.width=10, fig.height=5}
# Rank individual weight
SortWt <- lapply(hab, FUN=function(x)cbind(rank=1:length(x$Wt), weight=sort(x$Wt, decreasing=TRUE)))
SortWt <- ldply(SortWt)
info <- ldply(strsplit(SortWt$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat")
SortWt <- cbind(info, SortWt[, -1])

ggplot(data=SortWt, aes(x=weight, y=rank, shape=Habitat, colour=Habitat))+
  geom_point(size=2)+
  scale_shape_manual(values=c(19,1))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.LC), linetype=2)+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.UC), linetype=2)+
  facet_wrap(~Cruise, scales="free")+
  scale_x_log10(expression(Individual~weight~(mg)), labels = function(x) format(x, scientific = T))+
  scale_y_log10("Individual rank", limits=c(1, NA), labels = function(x) format(x, scientific = T))+
  theme_bw() %+replace% large %+replace% dark
```

# NBSS and ISD by cruise and depth strata (or station)

```{r, fig.width=10, fig.height=5}
hab <- splitBy(~Cruise+Habitat+Depth.zone, data=si)
wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=LBNbiom.method)
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat", "Depth")
nbss <- cbind(info, nbss[, -1])

#wt <- wt[unlist(lapply(wt, length))>9]
out <- lapply(wt, FUN=MLE.method)
# size specutrum slope estimate by MLE
lapply(out, FUN=function(x)x[1:2])
plb <- ldply(lapply(out, FUN=function(x)x$PLB))
info <- ldply(strsplit(plb$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat", "Depth")
plb <- cbind(info, plb[, -1])

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, shape=Habitat, colour=Habitat))+
  geom_point(size=3)+
  geom_path()+
  scale_shape_manual(values=c(19,1))+
  #geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB)))+
  #geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.LC)), linetype=2)+
  #geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.UC)), linetype=2)+
  facet_grid(Cruise~Depth)+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large %+replace% dark

# Only use size > 1e-3

hab <- splitBy(~Cruise+Habitat+Depth.zone, data=subset(si, Size>5e-4))
wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=LBNbiom.method)
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat", "Depth")
nbss <- cbind(info, nbss[, -1])

#wt <- wt[unlist(lapply(wt, length))>9]
out <- lapply(wt, FUN=MLE.method)
# size specutrum slope estimate by MLE
lapply(out, FUN=function(x)x[1:2])
plb <- ldply(lapply(out, FUN=function(x)x$PLB))
info <- ldply(strsplit(plb$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat", "Depth")
plb <- cbind(info, plb[, -1])

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, shape=Habitat, colour=Habitat))+
  geom_point(size=3)+
  scale_shape_manual(values=c(19,1))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB)))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.LC)), linetype=2)+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.UC)), linetype=2)+
  facet_grid(Cruise~Depth)+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large %+replace% dark
```

```{r, fig.width=10, fig.height=5}
# Rank individual weight
SortWt <- lapply(hab, FUN=function(x)cbind(rank=1:length(x$Wt), weight=sort(x$Wt, decreasing=TRUE)))
SortWt <- ldply(SortWt)
info <- ldply(strsplit(SortWt$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat", "Depth")
SortWt <- cbind(info, SortWt[, -1])

ggplot(data=SortWt, aes(x=weight, y=rank, shape=Habitat, colour=Habitat))+
  geom_point(size=2)+
  scale_shape_manual(values=c(19,1))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.LC), linetype=2)+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.UC), linetype=2)+
  facet_grid(Cruise~Depth)+
  scale_x_log10(expression(Individual~weight~(mg)), labels = function(x) format(x, scientific = T))+
  scale_y_log10("Individual rank", limits=c(1, NA), labels = function(x) format(x, scientific = T))+
  theme_bw() %+replace% large %+replace% dark
```

# NBSS and ISD by habitat and dominant taxa

```{r, fig.width=10, fig.height=5}
hab_pol <- splitBy(~Cruise+Habitat+Category, data=subset(si, Category=="Polychaeta" & (Cruise=="OR1_1096"|Cruise=="OR1_1114"|Cruise=="OR1_1126")))
hab_nem <- splitBy(~Cruise+Habitat+Category, data=subset(si, Category=="Nematoda" &  (Cruise=="OR1_1126"|Cruise=="OR1_1126")))
hab_har <- splitBy(~Cruise+Habitat+Category, data=subset(si, Category=="Harpacticoida" & (Cruise=="OR1_1096"|Cruise=="OR1_1102"|Cruise=="OR1_1114"|Cruise=="OR1_1126")))
hab <- rbind(ldply(hab_nem), ldply(hab_har), ldply(hab_pol))
hab <- splitBy(~Habitat+Category, hab)

wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=LBNbiom.method)
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- c("Habitat", "Category")
nbss <- cbind(info, nbss[, -1])

wt <- wt[unlist(lapply(wt, length))>21]
out <- lapply(wt, FUN=MLE.method)
# size specutrum slope estimate by MLE
lapply(out, FUN=function(x)x[1:2])
plb <- ldply(lapply(out, FUN=function(x)x$PLB))
info <- ldply(strsplit(plb$.id, split="[|]"))
names(info) <- c("Habitat", "Category")
plb <- cbind(info, plb[, -1])

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, shape=Habitat, colour=Habitat))+
  geom_point(size=3)+
  geom_path()+
  scale_shape_manual(values=c(19,1))+
  #geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB)))+
  #geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.LC)), linetype=2)+
  #geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.UC)), linetype=2)+
  facet_wrap(~Category, scales="free")+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large %+replace% dark

# Only use size > 1e-3

hab_pol <- splitBy(~Cruise+Habitat+Category, data=subset(si, Category=="Polychaeta" & (Cruise=="OR1_1096"|Cruise=="OR1_1114"|Cruise=="OR1_1126")&Size>1e-2))
hab_nem <- splitBy(~Cruise+Habitat+Category, data=subset(si, Category=="Nematoda" &  (Cruise=="OR1_1126"|Cruise=="OR1_1126")&Size>1e-4))
hab_har <- splitBy(~Cruise+Habitat+Category, data=subset(si, Category=="Harpacticoida" & (Cruise=="OR1_1096"|Cruise=="OR1_1102"|Cruise=="OR1_1114"|Cruise=="OR1_1126")&Size>1e-3))
hab <- rbind(ldply(hab_nem), ldply(hab_har), ldply(hab_pol))
hab <- splitBy(~Habitat+Category, hab)

wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=LBNbiom.method)
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- c("Habitat", "Category")
nbss <- cbind(info, nbss[, -1])

wt <- wt[unlist(lapply(wt, length))>21]
out <- lapply(wt, FUN=MLE.method)
# size specutrum slope estimate by MLE
lapply(out, FUN=function(x)x[1:2])
plb <- ldply(lapply(out, FUN=function(x)x$PLB))
info <- ldply(strsplit(plb$.id, split="[|]"))
names(info) <- c("Habitat", "Category")
plb <- cbind(info, plb[, -1])

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, shape=Habitat, colour=Habitat))+
  geom_point(size=3)+
  scale_shape_manual(values=c(19,1))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB)))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.LC)), linetype=2)+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.UC)), linetype=2)+
  facet_wrap(~Category, scales="free")+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large %+replace% dark
```

```{r, fig.width=10, fig.height=5}
# Rank individual weight
SortWt <- lapply(hab, FUN=function(x)cbind(rank=1:length(x$Wt), weight=sort(x$Wt, decreasing=TRUE)))
SortWt <- ldply(SortWt)
info <- ldply(strsplit(SortWt$.id, split="[|]"))
names(info) <- c("Habitat", "Category")
SortWt <- cbind(info, SortWt[, -1])

ggplot(data=SortWt, aes(x=weight, y=rank, shape=Habitat, colour=Habitat))+
  geom_point(size=2)+
  scale_shape_manual(values=c(19,1))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.LC), linetype=2)+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.UC), linetype=2)+
  facet_wrap(~Category, scales="free_x")+
  scale_x_log10(expression(Individual~weight~(mg)), labels = function(x) format(x, scientific = T))+
  scale_y_log10("Individual rank", limits=c(1, NA), labels = function(x) format(x, scientific = T))+
  theme_bw() %+replace% large %+replace% dark
```
