---
title: "Merged meiofauna and Macrofauna Size Spectrum"
author: "Chih-Lin Wei"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Meiofauna and macrofauna size from OR1 1114 and 1126 were merged and analyzed together.

```{r}
library(bbbs)
library(ggplot2)
library(plyr)
library(doBy)
#library(devtools)
#install_github("andrew-edwards/sizeSpectra")
library(sizeSpectra)

large <- theme(legend.title = element_text(size=20),
        legend.text = element_text(size=20),
        strip.text=element_text(size=20),
        axis.title = element_text(size=20),
        axis.text.y = element_text(size=20),
        axis.text.x = element_text(size=20, angle=60, hjust=0.5))

rotate <- theme(axis.text.x = element_text(size=20, angle=60, hjust=0.5))

# dark theme for ggplot
dark <- theme(plot.background = element_rect(colour = 'NA', fill = 'gray10'),
        panel.background = element_rect(colour = 'NA', fill = 'transparent'),
        legend.background = element_rect(colour = 'NA', fill = 'transparent'),
        legend.title = element_text(colour = 'white', size=20),
        legend.text = element_text(colour = 'white', size=20),
        axis.title = element_text(colour = 'white', size=20),
        axis.text = element_text(colour = 'white', size=20),
        axis.ticks = element_line(colour = 'white'),
        panel.border = element_rect(fill = 'NA', colour = 'white'), 
        panel.grid.major = element_line(colour = 'gray30'),
        panel.grid.minor = element_line(colour = 'gray20'))
# log2 bin
l2b <- function(x) 2^( floor(log2(min(x))) : ceiling(log2(max(x))) )
```


# Construction of meiofauna + macrofauna size specturm

Normalized biomass size spectrum (NBSS) and individual size distribution (ISD) were computed using R package sizeSpectra (Edwards et al., 2017, Methods in Ecology and Evolution 1:57-67). A probability density function based on a bounded power law (PLB) distribution was fitted to the ISD of meiofauna+macrofauna using maximum likelihood estimation (MLE) to estimate the exponent b. The exponent b (determined by MLE) is equivalent to the regression slope of the NBSS; however, comparison studies by Edwards et al. (2017) concluded that MLE method is the most accurate way to estimate the slope.
For normalized biomass size spectrum (NBSS), individual size of meiofauna+macrofauna were binned by log10 transformation of their biomass. The total biomass of each bin was normalized by the width of the bin. The log10 normalized biomass was plotted against the log10 midpoint of the size bin; however, the regression slope (and line) was not estimated from the NBSS. Instead, the slope (and line) was estimated by a customized probability density function using maximum likelihood.
The same fitted line (with 95% confidence interval) was plotted on the individual size distribution, which ranks the body size in decreasing order to visualize the fit between the model and size data.

# NBSS and ISD by habitat
## All size classes

```{r, fig.width=7, fig.height=5}
# Use functions from Edwards etal. (2017) Methods in Ecology and Evolution 1:57-67
# https://github.com/andrew-edwards/fitting-size-spectra
#source("PLBfunctions.r")

hab <- splitBy(~Habitat, data=si)
wt <- lapply(hab, FUN=function(x)x$Wt)
bks <- l2b(unlist(wt))
out <- lapply(wt, FUN=function(x)LBNbiom.method(x, binBreaks=bks))
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- "Habitat"
nbss <- cbind(info, nbss[, -1])

ggplot(data=ldply(hab), 
       aes(x=log10(Wt), fill=Habitat))+
  geom_density(alpha=0.5)+
  scale_fill_manual(values=c("black", "gray"))+
  labs(y="Frequency", x=expression(Log[10]~body~size~(mg)))+
  theme_bw() %+replace% large #%+replace% dark

ggplot(data=nbss,aes(x=log10binMid, y=log10totalBiom, colour=Habitat))+
  geom_point(size=5)+
  geom_path(size=1)+
  scale_colour_manual(values=c("black","gray"))+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~biomass~(mg~m^-2)))+
  theme_bw() %+replace% large #%+replace% dark

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, colour=Habitat))+
  geom_point(size=3)+
  geom_path(size=1)+
  scale_colour_manual(values=c("black", "gray"))+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large #%+replace% dark
```

# Size classes >= 1e-03

```{r, fig.width=7, fig.height=5}
# Only use Wt>= 1e-03

hab <- splitBy(~Habitat, data=subset(si, Wt>=1e-03))
wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=function(x)LBNbiom.method(x, binBreaks=bks))
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- "Habitat"
nbss <- cbind(info, nbss[, -1])

#wt <- wt[unlist(lapply(wt, length))>9]
out <- lapply(wt, FUN=MLE.method)
# size specutrum slope estimate by MLE
lapply(out, FUN=function(x)x[1:2])
plb <- ldply(lapply(out, FUN=function(x)x$PLB))
info <- ldply(strsplit(plb$.id, split="[|]"))
names(info) <- "Habitat"
plb <- cbind(info, plb[, -1])

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, colour=Habitat))+
  geom_point(size=3)+
  scale_colour_manual(values=c("black", "gray"))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB)))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.LC)), linetype=2)+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.UC)), linetype=2)+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large #%+replace% dark

# Rank individual weight
SortWt <- lapply(hab, FUN=function(x)cbind(rank=1:length(x$Wt), weight=sort(x$Wt, decreasing=TRUE)))
SortWt <- ldply(SortWt)
info <- ldply(strsplit(SortWt$.id, split="[|]"))
names(info) <- "Habitat"
SortWt <- cbind(info, SortWt[, -1])

ggplot(data=SortWt, aes(x=weight, y=rank, colour=Habitat))+
  geom_point(size=2)+
  scale_colour_manual(values=c("black", "gray"))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.LC), linetype=2)+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.UC), linetype=2)+
  #facet_wrap(~Cruise, scales="free")+
  scale_x_log10(expression(Individual~weight~(mg)), labels = function(x) format(x, scientific = T))+
  scale_y_log10("Individual rank", limits=c(1, NA), labels = function(x) format(x, scientific = T))+
  theme_bw() %+replace% large #%+replace% dark
```

# NBSS and ISD by cruise and habitat
## All size classes

```{r, fig.width=7, fig.height=7}
hab <- splitBy(~Cruise+Habitat, data=si)
wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=function(x)LBNbiom.method(x, binBreaks=bks))
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat")
nbss <- cbind(info, nbss[, -1])

ggplot(data=ldply(hab), 
       aes(x=log10(Wt), fill=Habitat))+
  geom_density(alpha=0.5)+
  scale_fill_manual(values=c("black", "gray"))+
  facet_wrap(~Cruise, scales="free")+
  labs(y="Frequency", x=expression(Log[10]~body~size~(mg)))+
  theme_bw() %+replace% large #%+replace% dark

ggplot(data=nbss,aes(x=log10binMid, y=log10totalBiom, colour=Habitat))+
  geom_point(size=5)+
  geom_path(size=1)+
  facet_wrap(~Cruise, scales="free")+
  scale_colour_manual(values=c("black","gray"))+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~biomass~(mg~m^-2)))+
  theme_bw() %+replace% large #%+replace% dark

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, colour=Habitat))+
  geom_point(size=3)+
  geom_path()+
  scale_colour_manual(values=c("black", "gray"))+
  facet_wrap(~Cruise)+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large #%+replace% dark
```

## Size classes >= 1e-03

```{r, fig.width=10, fig.height=10}
# Only use Wt> 1e-03

hab <- splitBy(~Cruise+Habitat, data=subset(si, Wt>=1e-03))
wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=function(x)LBNbiom.method(x, binBreaks=bks))
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat")
nbss <- cbind(info, nbss[, -1])

#wt <- wt[unlist(lapply(wt, length))>9]
out <- lapply(wt, FUN=MLE.method)
# size specutrum slope estimate by MLE
lapply(out, FUN=function(x)x[1:2])
plb <- ldply(lapply(out, FUN=function(x)x$PLB))
info <- ldply(strsplit(plb$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat")
plb <- cbind(info, plb[, -1])

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, colour=Habitat))+
  geom_point(size=3)+
  scale_colour_manual(values=c("black", "gray"))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB)))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.LC)), linetype=2)+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.UC)), linetype=2)+
  facet_wrap(~Cruise)+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large #%+replace% dark

# Rank individual weight
SortWt <- lapply(hab, FUN=function(x)cbind(rank=1:length(x$Wt), weight=sort(x$Wt, decreasing=TRUE)))
SortWt <- ldply(SortWt)
info <- ldply(strsplit(SortWt$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat")
SortWt <- cbind(info, SortWt[, -1])

ggplot(data=SortWt, aes(x=weight, y=rank, colour=Habitat))+
  geom_point(size=2)+
  scale_colour_manual(values=c("black", "gray"))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.LC), linetype=2)+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.UC), linetype=2)+
  facet_wrap(~Cruise, scales="free")+
  scale_x_log10(expression(Individual~weight~(mg)), labels = function(x) format(x, scientific = T))+
  scale_y_log10("Individual rank", limits=c(1, NA), labels = function(x) format(x, scientific = T))+
  theme_bw() %+replace% large #%+replace% dark
```

# NBSS and ISD by depth strata and habitat
## All size classes

```{r, fig.width=10, fig.height=5}
hab <- splitBy(~Habitat+Depth.zone, data=si)
wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=function(x)LBNbiom.method(x, binBreaks=bks))
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- c("Habitat", "Depth")
nbss <- cbind(info, nbss[, -1])

ggplot(data=ldply(hab), 
       aes(x=log10(Wt), fill=Habitat))+
  geom_density(alpha=0.5)+
  scale_fill_manual(values=c("black", "gray"))+
  facet_wrap(~Depth.zone, scales="free")+
  labs(y="Frequency", x=expression(Log[10]~body~size~(mg)))+
  theme_bw() %+replace% large #%+replace% dark

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiom, colour=Habitat))+
  geom_point(size=3)+
  geom_path(size=1)+
  scale_colour_manual(values=c("black","gray"))+
  facet_wrap(~Depth, scales="free")+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~biomass~(mg~m^-2)))+
  theme_bw()%+replace% large #%+replace% dark

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, colour=Habitat))+
  geom_point(size=3)+
  geom_path()+
  scale_colour_manual(values=c("black", "gray"))+
  facet_wrap(~Depth, scales="free")+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large #%+replace% dark
```

## Size classes >= 1e-03

```{r, fig.width=10, fig.height=5}
# Only use Wt>= 1e-03

hab <- splitBy(~Habitat+Depth.zone, data=subset(si, Wt>= 1e-03))
wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=function(x)LBNbiom.method(x, binBreaks=bks))
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- c("Habitat", "Depth")
nbss <- cbind(info, nbss[, -1])

#wt <- wt[unlist(lapply(wt, length))>9]
out <- lapply(wt, FUN=MLE.method)
# size specutrum slope estimate by MLE
lapply(out, FUN=function(x)x[1:2])
plb <- ldply(lapply(out, FUN=function(x)x$PLB))
info <- ldply(strsplit(plb$.id, split="[|]"))
names(info) <- c("Habitat", "Depth")
plb <- cbind(info, plb[, -1])

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, colour=Habitat))+
  geom_point(size=3)+
  scale_colour_manual(values=c("black", "gray"))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB)))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.LC)), linetype=2)+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.UC)), linetype=2)+
  facet_wrap(~Depth, scales="free")+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large #%+replace% dark

# Rank individual weight
SortWt <- lapply(hab, FUN=function(x)cbind(rank=1:length(x$Wt), weight=sort(x$Wt, decreasing=TRUE)))
SortWt <- ldply(SortWt)
info <- ldply(strsplit(SortWt$.id, split="[|]"))
names(info) <- c("Habitat", "Depth")
SortWt <- cbind(info, SortWt[, -1])

ggplot(data=SortWt, aes(x=weight, y=rank, colour=Habitat))+
  geom_point(size=2)+
  scale_colour_manual(values=c("black", "gray"))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.LC), linetype=2)+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.UC), linetype=2)+
  facet_wrap(~Depth, scales="free")+
  scale_x_log10(expression(Individual~weight~(mg)), labels = function(x) format(x, scientific = T))+
  scale_y_log10("Individual rank", limits=c(1, NA), labels = function(x) format(x, scientific = T))+
  theme_bw() %+replace% large #%+replace% dark
```

# NBSS and ISD by cruise and depth strata (or station)
## All size classes

```{r, fig.width=10, fig.height=10}
hab <- splitBy(~Cruise+Habitat+Depth.zone, data=si)
wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=function(x)LBNbiom.method(x, binBreaks=bks))
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat", "Depth")
nbss <- cbind(info, nbss[, -1])

ggplot(data=ldply(hab), 
       aes(x=log10(Wt), fill=Habitat))+
  geom_density(alpha=0.5)+
  scale_fill_manual(values=c("black", "gray"))+
  facet_grid(Cruise~Depth.zone, scales="free")+
  labs(y="Frequency", x=expression(Log[10]~body~size~(mg)))+
  theme_bw() %+replace% large #%+replace% dark

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiom, colour=Habitat))+
  geom_point(size=3)+
  geom_path(size=1)+
  scale_colour_manual(values=c("black","gray"))+
  facet_grid(Cruise~Depth, scales="free")+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~biomass~(mg~m^-2)))+
  theme_bw()%+replace% large #%+replace% dark

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, colour=Habitat))+
  geom_point(size=3)+
  geom_path()+
  scale_colour_manual(values=c("black", "gray"))+
  facet_grid(Cruise~Depth, scales="free")+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large #%+replace% dark
```

## Size classes >= 1e-03

```{r, fig.width=10, fig.height=10}
# Only use Wt>= 1e-03

hab <- splitBy(~Cruise+Habitat+Depth.zone, data=subset(si, Wt>= 1e-03))
wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=function(x)LBNbiom.method(x, binBreaks=bks))
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat", "Depth")
nbss <- cbind(info, nbss[, -1])

wt <- wt[unlist(lapply(wt, length))>174]
out <- lapply(wt, FUN=MLE.method)
# size specutrum slope estimate by MLE
lapply(out, FUN=function(x)x[1:2])
plb <- ldply(lapply(out, FUN=function(x)x$PLB))
info <- ldply(strsplit(plb$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat", "Depth")
plb <- cbind(info, plb[, -1])

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, colour=Habitat))+
  geom_point(size=3)+
  scale_colour_manual(values=c("black", "gray"))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB)))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.LC)), linetype=2)+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.UC)), linetype=2)+
  facet_grid(Cruise~Depth, scales="free")+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large #%+replace% dark

# Rank individual weight
SortWt <- lapply(hab, FUN=function(x)cbind(rank=1:length(x$Wt), weight=sort(x$Wt, decreasing=TRUE)))
SortWt <- ldply(SortWt)
info <- ldply(strsplit(SortWt$.id, split="[|]"))
names(info) <- c("Cruise", "Habitat", "Depth")
SortWt <- cbind(info, SortWt[, -1])

ggplot(data=SortWt, aes(x=weight, y=rank, colour=Habitat))+
  geom_point(size=2)+
  scale_colour_manual(values=c("black", "gray"))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.LC), linetype=2)+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.UC), linetype=2)+
  facet_grid(Cruise~Depth, scales="free")+
  scale_x_log10(expression(Individual~weight~(mg)), labels = function(x) format(x, scientific = T))+
  scale_y_log10("Individual rank", limits=c(1, NA), labels = function(x) format(x, scientific = T))+
  theme_bw() %+replace% large #%+replace% dark
```

# NBSS and ISD by habitat and dominant taxa
## All size classes

```{r, fig.width=10, fig.height=5}
hab_pol <- splitBy(~Cruise+Habitat+Category, data=subset(si, Category=="Polychaeta" & (Cruise=="OR1_1096"|Cruise=="OR1_1114"|Cruise=="OR1_1126")))
hab_nem <- splitBy(~Cruise+Habitat+Category, data=subset(si, Category=="Nematoda" &  (Cruise=="OR1_1114"|Cruise=="OR1_1126")))
hab_har <- splitBy(~Cruise+Habitat+Category, data=subset(si, Category=="Harpacticoida" & (Cruise=="OR1_1096"|Cruise=="OR1_1102"|Cruise=="OR1_1114"|Cruise=="OR1_1126")))
hab <- rbind(ldply(hab_nem), ldply(hab_har), ldply(hab_pol))
hab <- splitBy(~Habitat+Category, hab)

wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=function(x)LBNbiom.method(x, binBreaks=bks))
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- c("Habitat", "Category")
nbss <- cbind(info, nbss[, -1])

#wt <- wt[unlist(lapply(wt, length))>21]
out <- lapply(wt, FUN=MLE.method)

ggplot(data=ldply(hab), 
       aes(x=log10(Wt), fill=Habitat))+
  geom_density(alpha=0.5)+
  scale_fill_manual(values=c("black", "gray"))+
  facet_grid(Cruise~Category, scales="free")+
  labs(y="Frequency", x=expression(Log[10]~body~size~(mg)))+
  theme_bw() %+replace% large #%+replace% dark

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiom, colour=Habitat))+
  geom_point(size=3)+
  geom_path(size=1)+
  scale_colour_manual(values=c("black","gray"))+
  facet_wrap(~Category, scales="free")+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~biomass~(mg~m^-2)))+
  theme_bw()  %+replace% large  %+replace% rotate #%+replace% dark

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, colour=Habitat))+
  geom_point(size=3)+
  geom_path(size=1)+
  scale_colour_manual(values=c("black", "gray"))+
  facet_wrap(~Category, scales="free")+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large #%+replace% dark
```

```{r, fig.width=10, fig.height=5}
# Only use Wt> 9.765625e-04 (harpacticoids), 9.765625e-04 (nematods), and 1.562500e-02 (polychaetes)

hab_pol <- splitBy(~Cruise+Habitat+Category, data=subset(si, Category=="Polychaeta" & (Cruise=="OR1_1096"|Cruise=="OR1_1114"|Cruise=="OR1_1126")&Wt>1e-03))
hab_nem <- splitBy(~Cruise+Habitat+Category, data=subset(si, Category=="Nematoda" &  (Cruise=="OR1_1126"|Cruise=="OR1_1126")&Wt>1e-03))
hab_har <- splitBy(~Cruise+Habitat+Category, data=subset(si, Category=="Harpacticoida" & (Cruise=="OR1_1096"|Cruise=="OR1_1102"|Cruise=="OR1_1114"|Cruise=="OR1_1126")&Wt>1e-02))
hab <- rbind(ldply(hab_nem), ldply(hab_har), ldply(hab_pol))
hab <- splitBy(~Habitat+Category, hab)

wt <- lapply(hab, FUN=function(x)x$Wt)
out <- lapply(wt, FUN=function(x)LBNbiom.method(x, binBreaks=bks))
nbss <- ldply(lapply(out, FUN=function(x)x$binVals), data.frame)
info <- ldply(strsplit(nbss$.id, split="[|]"))
names(info) <- c("Habitat", "Category")
nbss <- cbind(info, nbss[, -1])

#wt <- wt[unlist(lapply(wt, length))>21]
out <- lapply(wt, FUN=MLE.method)
# size specutrum slope estimate by MLE
lapply(out, FUN=function(x)x[1:2])
plb <- ldply(lapply(out, FUN=function(x)x$PLB))
info <- ldply(strsplit(plb$.id, split="[|]"))
names(info) <- c("Habitat", "Category")
plb <- cbind(info, plb[, -1])

ggplot(data=nbss, aes(x=log10binMid, y=log10totalBiomNorm, colour=Habitat))+
  geom_point(size=3)+
  scale_colour_manual(values=c("black", "gray"))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB)))+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.LC)), linetype=2)+
  geom_path(data=plb, aes(x=log10(x.PLB), y=log10(B.PLB.UC)), linetype=2)+
  facet_wrap(~Category, scales="free")+
  labs(x=expression(Log[10]~"mid-point of"~size~class~(mg)), y=expression(Log[10]~normalized~biomass))+
  theme_bw() %+replace% large #%+replace% dark

# Rank individual weight
SortWt <- lapply(hab, FUN=function(x)cbind(rank=1:length(x$Wt), weight=sort(x$Wt, decreasing=TRUE)))
SortWt <- ldply(SortWt)
info <- ldply(strsplit(SortWt$.id, split="[|]"))
names(info) <- c("Habitat", "Category")
SortWt <- cbind(info, SortWt[, -1])

ggplot(data=SortWt, aes(x=weight, y=rank, colour=Habitat))+
  geom_point(size=2)+
  scale_colour_manual(values=c("black", "gray"))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB))+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.LC), linetype=2)+
  geom_path(data=plb, aes(x=x.PLB, y=y.PLB.UC), linetype=2)+
  facet_wrap(~Category, scales="free_x")+
  scale_x_log10(expression(Individual~weight~(mg)), labels = function(x) format(x, scientific = T))+
  scale_y_log10("Individual rank", limits=c(1, NA), labels = function(x) format(x, scientific = T))+
  theme_bw() %+replace% large #%+replace% dark
```
